---
title: 'Personal R notes'
output: github_document
---
  
# `rmarkdown`
## Add a comment
```{r}
[//]: # "Comment”
  [//]: # (Comment)
  ```

## Common mistakes
+ Two chunks with the same name will cause a big error message   
+ Bullet list need to have an empty line before to be recognized it can also be forced with html code `<li>item1</li>` or \LaTeX  `\list{\item \item}`

## CSS styling in HTML
```{r}
# Centering knitr::kable() on the page
<style type="text/css">
  .table {
    margin-left:30%; 
    width:40%;
  }
</style>
  ```

## Template settings for \LaTeX
```{r}
---
  title: 
  author: "Benjamin Guiastrennec"
date: "`r format(Sys.time(), '%d %B, %Y')`"
mainfont: Arial
sansfont: Arial
monofont: Courier
output:
  pdf_document:
  latex_engine: xelatex
number_sections: yes
toc: yes
---
  ```

## Change `knitr` options
```{r}
# Change default to eval = FALSE
knitr::opts_chunk$set(eval = FALSE)

# Enable soft wrap in latex
knitr::opts_chunk$set(tidy = TRUE, tidy.opts = list(width.cutoff = 60))
```
*Note : see also opt_knit$set() for changing global knitr options*
  
  ## Extract Rmarkdown R code
  ```{r}
knitr::purl('file.Rmd', documentation = 2)
```

## Chunk options
```{r}
tidy = TRUE   # Wrap code
```
