---
title: 'Personal R notes'
output: github_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(eval = FALSE) # Change knitr defaults
```

# `ggplot2`
## Use special symbol in labels
+ Text labels: `parse = T` inside `geom_text()` or `annotate()`
+ Axis labels: `expression(alpha)` to get greek alpha
+ Facet labels: Use `labeller = label_parsed()` inside `facet`
+ Legend labels: Use `bquote(alpha == .(value))`

Note:

+ `bold('bold')`, `italic('italic')`, `atop('top','bottom')`
+ `*` is a non space separator, `~` is to add spaces
+ `[]` for subscript and `^` for superscript
+ `bquote()` when substitutions are needed `.(sub)` to substitute an object
+ `help('plotmath')` for more info

Example: `labs(x = expression(bold(T[max])~(h))`

## Transparent rectangles
Add transparent rectangle to highlight graph area
```{r}
# Create data for rectangle
rect <- data.frame(xmin = 0, xmax = 4, ymin = -Inf, ymax = Inf)

# Create plot
ggplot(data = data, aes(x =, y =))+
geom_rect(data = rect, aes(xmin = xmin, xmax = xmax, 
                           ymin = ymin, ymax = ymax),
          fill = 'steelblue3', alpha = 0.1, inherit.aes = FALSE)
```

## Add mean or median to dotplot
```{r}
library(ggplot2)
ggplot(data = data, aes(x = x, y = y, color = color)) + geom_point() +
geom_errorbar(stat = 'hline', yintercept = 'mean', width = 0.8, 
              aes(ymax = ..y.., ymin = ..y..))
```

## Scale `geom_density()`
```{r}
ggplot(data, aes(x = x))+
  geom_histogram(aes(y = ..density..))+
  geom_density()
```
_Note: `..scaled..` (0-1) and `..count..` can also be used_

## Correlation graphs
```{r} 
library(GGally)
ggpairs(data, columns = 1:7, colour = 'sex', 
        lower = list(continuous = 'smooth'), 
        diag  = list(continuous = 'density', discrete = 'bar'),
        axisLabels = 'show')
```

## `gridExtra`
`grid.arrange()` draws directly on a device while `arrangeGrob()` doesn't draw anything but returns a grob
```{r}
# Plot 4 graphs with different panel size
grid.arrange(P1, P2, P3, P4, 
             ncol = 2, nrow = 2, 
             widths = c(4, 1), heights = c(1, 4))
```

## Special smooth
```{r}
# 90% CI of the Independent variable
CI90     <- quantile(Sim$TIME, c(0.05, 0.95))

# Generate ggplot with smooth
Gsmooth  <- ggplot(data = Sim, aes(x = TIME, y = WRES))+
              stat_smooth(method = 'loess', se = FALSE)

# Get the data from the smooth
Gsmooth  <- as.data.frame(ggplot_build(Gsmooth)$data)

# Subset the 90% CI of the smooth
Gsmooth  <- Gsmooth[Gsmooth$x >= CI90[1] & Gsmooth$x <= CI90[2],]

ggplot()+
  
  # Rectangle showing the 90%CI
  geom_rect(aes(xmin = CI90[1], xmax = CI90[2],
                ymax = Inf, ymin = -Inf), 
           alpha = 0.2, fill = 'dodgerblue3')+
  
  # Full smooth as a grey dashed line
  geom_smooth(data = Sim, aes(x = TIME,y = WRES), 
              method = 'loess', se = FALSE, size = 1, 
              col = 'grey70', linetype = 2)+
  
  # 90% CI smooth as black continuous line
  geom_line(data = Gsmooth, aes(x = x, y = y), size = 1, col = 'black')+
  
  # Add actual data
  geom_point(data = Sim, aes(x = TIME, y = WRES), col = 'grey35')+
  
  # Add labels
  geom_text(aes(x = mean(CI90), y = -6.5), 
            label = '90% confidence interval',
            col = 'dodgerblue4', alpha = 0.7, size = 4)
```

## Scales
### New color schemes (+ invert color scale)
```{r}
scale_fill_manual(values = rev(RColorBrewer::brewer.pal(6, 'Reds')))
```

### Remove legend for a given aesthetic 
```{r}
scale_linetype_identity()
```

### Remove an axis
```{r}
scale_x_continuous(NULL, breaks = NULL)
```

### Expand the limits
```{r}
expand_limits()
```

### Special breaks for each panel
```{r}
ebreaks <- function(...) {
  function(x) { # x contains min and max for each panel
    if(x[1] < 120) { breaks <- seq(0, 200, 10)
    } else { breaks <- seq(150, 350, 50) }
    return(breaks) # Return all breaks
}}
ggplot(mtcars, aes(x=hp, y = disp))+ geom_point()+
  scale_x_continuous(breaks = ebreaks())+
  facet_wrap(~cyl, scales = 'free_x')
```


## Annotations
### Based on graph coordinates
```{r}
annotate(..., nudge_y = 3, nudge_x = 10)
```

### Based on relative positioning
```{r}
annotate(..., hjust = 1 , vjust = 0)
```

### Special annotations
```{r}
annotation_custom()    # Special annotation function
annotation_logticks()  # For creating nice logscale
```


## Control the legend
```{r}
guides(fill = guide_legend(reverse = TRUE))
```
See also `guide_colorbars()`


## Access the `ggplot2 lm()` data
```{r}
fortify.lm()
```


## Use the labeller function
```{r}
ggplot(mtcars, aes(x = disp, y = mpg)) + 
  geom_point() + 
  facet_wrap(~ vs + am, 
             label = labeller(.default = label_both, .multi_line = FALSE)) 
```


## Sliders with manipulate
```{r}
library(manipulate)
manipulate(
 ggplot(data = data, aes(x = , y = ))+ geom_line(),
 prm = slider(0, 20, 10, step = 1)
)
```
